---
title: "Meta-Analysis for Longevity Summary Excluding HUM251"
author: "Fay Frost"
output: pdf_document
---
  
```{r, global_options, include=FALSE}
knitr::opts_chunk$set(message=FALSE, tidy.opts=list(width.cutoff=60), tidy=TRUE) 

pacman::p_load(here, ape, phytools, metafor, orchaRd, ggtree)

```

# 1. Summary

This document reports the process taken in the model fitting stage of the meta-analysis in thermal longevity. 

# 2. Setup
We first read in our data and select all of the effect sizes related to longevity. We do this using the following code.

```{r, eval=TRUE, include=FALSE}
### Read in effect size data
effectdata <- read.csv(here::here("data", "Survival project all pairwise.es.csv"))
longdata_warm <- subset(effectdata, Trait.category == "Longevity" & warm.cool == "Warm" )
longdata_cool <- subset(effectdata, Trait.category == "Longevity" & warm.cool == "Cool" )

alllong <- rbind(longdata_warm, longdata_cool)

### select data for analysis
rdata <- alllong

rdata <- subset(rdata, Paper.code != "HUM251")
```

Next we create new columns in our dataframe which will serve as random factors in our multi-level meta analysis models. The following initialises four new columns, namely "obs", "study_code", "Species.phylo" and "species". Lastly, we create a column name "precision" which is equal to the inverse standard error. 

```{r, eval=TRUE, include=FALSE}
### Create random factors into data frame 
rdata$obs <- factor(c(1:nrow(rdata)))                # Unique observation code
rdata$study_code <- factor(rdata$Paper.code)         # Model requires column names study_code 
rdata$Species.phylo <- factor(rdata$Species.latin)   # Species names for phylo matrix
rdata$species <- factor(rdata$Species.latin)         # Another species column for random factor 

precision <- sqrt(1/rdata$v)                         # inverse standard error 
rdata[,"precision"] <- precision
```

The number of species and total number of studies present in the data are as follows.
```{r}
nlevels(rdata$species)    # Check number of species
nlevels(rdata$study_code) # Check number of studies
```

\newpage

The final stage in the setup is to import a phylogentic tree of the data. Below is the code used to produce the tree and a plot of the tree itself. 

```{r, eval=TRUE, include=FALSE}

## import tree from map
tree1 <- read.nexus(here::here("phylogeny","all_long_excHUM251_tree.nex"))
tree_grafen = compute.brlen(tree1, method="Grafen", power=1)
phylo_matrix <- vcv(tree_grafen, cor=TRUE, model="Brownian") # Make phylogenetic matrix
```

```{r, eval=TRUE, echo=FALSE}
###################

## import tree from map
notin <- setdiff(tree1$tip.label, rdata$Species.latin)
indata <- setdiff(tree1$tip.label, notin)

## prune tree
prune_tree <- drop.tip(tree_grafen, notin)
phylo_matrix <- vcv(prune_tree, cor=TRUE, model="Brownian")

## number of indata should equal # of unique rdata species. 
## we are currently 15 out.
setdiff(unique(rdata$Species.latin), indata)


##### create data frame and plot tree ###################
plot_data <- c()
plot_data$species_latin <- indata
plot_data$class <- rdata$Class[match(indata, rdata$Species.latin)]
plot_data <- as.data.frame(unique(plot_data))
colnames(plot_data) <- c("species_latin", "class")

plot_data$class[which(plot_data$class == "Arachnid")] <- "Arachnida"
plot_data$class[which(plot_data$class == "Insect")] <- "Insecta"

```


```{r, echo=FALSE}
ggtree(prune_tree, layout = "circular", lwd = 0.1) %<+% plot_data +
  geom_tiplab(size=1.2, offset=0.01) +
  aes(col = class) 
```

\newpage 

# 3. Random effects models

In this section we determine which random effects to include in our model. For each model I have provided the code used to specify the structure of the model and a summary of the results. We begin with a model that includes all of the random factors we created earlier. 

```{r, eval=FALSE}
# Adding four random factors
meta2 <- rma.mv(es, v, random= list(~ 1|Species.phylo, ~ 1|species, ~ 1|study_code, ~1|obs), 
                R= list(Species.phylo = phylo_matrix), data= rdata, method= "REML")
```

```{r}
summary(meta2)
i2_ml(meta2, method=c("ratio")) # Heterogeneity at each random factor level
```

## Accounting for non-independence of data points from the same experiment

The data has a nested structure. Each study (study_code) may have a number of experiments (effect.size.code) which share a common control temperature. Each effect size has its own unique code, obs. Effect sizes from the same experiment which share a control temperature are thought to be non-independent. The following code create a covariance matrix "VCV_shared" which assumes a correlation of 0.5 between effect sizes from the same experiment. We include this structure in our proceeding models.

```{r, eval=FALSE}
rdata$shared_control <- factor(rdata$Effect.size.code)
vcv_shared <- impute_covariance_matrix(vi=rdata$v, cluster = rdata$shared_control, r=0.5)

```

```{r, eval=FALSE}
# Add new variance matrix into the mixed-effects meta-analysis model
meta3 <- rma.mv(es, vcv_shared, random= list(~ 1|Species.phylo, ~ 1|species, ~ 1|study_code,  ~1|obs), R= list(Species.phylo = phylo_matrix), data= rdata, method= "REML")

```

```{r}
summary(meta3)
i2_ml(meta3, method=c("ratio")) # Heterogeneity at each random factor level
```

\newpage

## Model without phylogeny
The variance-covariance matrix for phylogenetic relatedness of included species has now been excluded as a random effect in the model (Chamberlain et al., 2012) as its inclusion did not improve
model fit and the phylogenetic signal was very weak.

```{r, eval=FALSE}
## without phylogeny but with shared control 
meta5 <- rma.mv(es, VCV_shared, random= list(~ 1|species, ~ 1|study_code,  ~1|obs), data= rdata, method= "REML")
```

```{r}
summary(meta5)
i2_ml(meta5, method=c("ratio")) # Heterogeneity at each random factor level
```

\newpage

## Model without phylogeny or species
```{r, eval=FALSE}
## without phylogeny or species
meta4 <- rma.mv(es, VCV_shared, random= list(~ 1|study_code,  ~1|obs), data= rdata, method= "REML")
```

```{r}
summary(meta4)
i2_ml(meta4, method=c("ratio")) # Heterogeneity at each random factor level
```

\newpage

## Model without phylogeny, species or study_code
```{r, eval=FALSE}
## without phylogeny, species or study_code 
meta7 <- rma.mv(es, VCV_shared, random= list(~1|obs), data= rdata, method= "REML")
```

```{r}
summary(meta7)
i2_ml(meta7, method=c("ratio")) # Heterogeneity at each random factor level
```

\newpage

## Model without phylogeny, species or shared_control 
```{r, eval=FALSE}
## without phylogeny, species or shared_control 
meta8 <- rma.mv(es, VCV_shared, random= list(~ 1|study_code, ~1|obs), data= rdata, method= "REML")
```

```{r}
summary(meta8)
i2_ml(meta8, method=c("ratio")) # Heterogeneity at each random factor level
```

We can see from the above that the best fitting model according to AIC is "meta8" which includes only the study code and the unique effect size code, obs. However, since there is not a huge difference ($\Delta$ AIC = 2) between the AIC of this model and "meta4" which also includes shared_control, we choose to proceed with meta4. The thought being that 

\newpage

# 4. Meta-regressions

Starting with the best fitting random-effect model from Section 3, "meta4" we now include single factors as a fixed effect. We initially explore the fixed factors

* **reftemp**: The experiment's control (reference) temperature.

* **treattemp**: The treatment temperature, which we expect to havea non-linear relationship to longevity.

* **warm.cool** : A categorical variable indicating whether treatment is warmer or cooler than the reference temperature 

* **diff**: The difference between the reference and treatment temperature.

## Reference temperature 
```{r, eval=FALSE}
meta_trait_ref <- rma.mv(es, VCV_shared, mod= ~reftemp, random= list(~ 1|study_code, ~1|obs), data= rdata, method= "REML")
```

```{r}
summary(meta_trait_ref)
```

\newpage

## Treatment temperature
```{r, eval=FALSE}
meta_trait_treattemp <- rma.mv(es, VCV_shared,  mod= ~treattemp, random= list(~ 1|study_code, ~1|obs), data= rdata, method= "REML")
```

```{r}
summary(meta_trait_treattemp)
```

\newpage

## Warm vs Cool
```{r, eval=FALSE}
meta_trait_warm <- rma.mv(es, VCV_shared,  mod= ~warm.cool, random= list(~ 1|study_code,  ~1|obs), data= rdata, method= "REML")
```

```{r}
summary(meta_trait_warm)
```

\newpage 

We model warm versus cool without and intercept so we can visualise the estimates easier. 

```{r, eval=FALSE}
meta_trait_warm_nointer <- rma.mv(es, VCV_shared,  mod= ~warm.cool-1, random= list(~ 1|study_code,  ~1|obs), data= rdata, method= "REML")
```

```{r}
summary(meta_trait_warm_nointer)
```

\newpage

## Difference
```{r, eval=FALSE}
meta_trait_diff <- rma.mv(es, VCV_shared,  mod= ~diff, random= list(~ 1|study_code,  ~1|obs), data= rdata, method= "REML")
```

```{r}
summary(meta_trait_diff)
```

\newpage
## Treatment temperature as a cubic effect

We expect that treatment temperature will have a non-linear effect on longevity. We expect that the relationship will be cubic. More specifically, we expect that longevity will increase at cooler temperature and decrease at more extreme higher temperature. We also assume that this optimum temperature will be close to the control temperature. That said, we investigate **treattemp** as a cubic fixed effect. 


```{r, eval=FALSE}
meta_trait_treat3 <- rma.mv(es, VCV_shared,  mod= ~ poly(treattemp, degree=3, raw=TRUE), random= list(~ 1|study_code, ~1|obs), data= rdata, method= "REML")
```

```{r}
summary(meta_trait_treat3)
```

For completeness we also investigate treatment temperature as a quadratic effect. 

```{r, eval=FALSE}
meta_trait_treat2 <- rma.mv(es, VCV_shared,  mod= ~ poly(treattemp, degree=2, raw=TRUE), random= list(~ 1|study_code,  ~1|obs), data= rdata, method= "REML")
```

```{r}
summary(meta_trait_treat2)
```

\newpage

Below is a bubble plot of the fitted cubic model. 

```{r, eval=TRUE, echo=FALSE}
p <- bubble_plot(meta_trait_treat3, mod = "treattemp", group="study_code")
p+coord_cartesian(ylim=c(-10,10))
```

\newpage 

## Modelling response with binned treatment temperatures

The last model we will try is one with categorised or 'binned' treatment temperatures. We categories our effect sizes into one seven bins depending on the treatment temperature. The bins and the number of effect sizes in each bin are given below


```{r}
table(rdata$bin.temp)
```

```{r, eval=FALSE, echo=FALSE}

# Create a custom order for bin.temp
rdata$bin.temp <- factor(rdata$bin.temp, levels = c("<15", "15-20", "20-25", "25-30", "30-35", "35-40", ">40"))

# Define custom colors for the boxplot
boxplot_colors <- c("<15" = "red", "15-20" = "blue", "20-25" = "green", "25-30" = "orange", "30-35" = "purple", "35-40" = "pink", ">40" = "brown")

# Create the boxplot using ggplot2 with custom colors and x-axis breaks
ggplot(rdata, aes(x = bin.temp, y = es, fill = bin.temp)) +
  geom_boxplot() +
  scale_fill_manual(values = boxplot_colors) +  # Apply custom colors
  labs(x = "Binned Temperature", y = "effect size") +
  ggtitle("Boxplot of effect sizes per temperature") +
  theme_minimal() +
  coord_cartesian(ylim = c(-10, 10))  # Set y-axis limits
```

```{r, eval=FALSE}
meta_trait_bintemp <- rma.mv(es, VCV_shared,  mod= ~bin.temp-1,  random= list(~ 1|study_code, ~1|obs), data= rdata, method= "REML")
```

```{r}
summary(meta_trait_bintemp)
```

```{r, echo=FALSE, eval=TRUE}
plot <- orchard_plot(meta_trait_bintemp, mod = "bin.temp", xlab = "Effect size", transfm = "none", 
             cb=FALSE, group="obs", tree.order = c("<15", "15-20", "20-25", "25-30", "30-35", "35-40", ">40")) 

plot+ coord_cartesian(ylim=c(-20,10))
```

\newpage

# Publication Bias.

We fit meta-regression model with year as a moderator to see whether there is an publication bias. In other words, is effect size strongly correlated with publication year. We found now evidence to suggest that year has an effect on the report effect sizes. A summary of the model is given below. 

```{r, eval=FALSE}
meta_year <- rma.mv(es, VCV_shared,  mod= ~Publication.year,  random= list(~ 1|study_code, ~1|obs), data= rdata, method= "REML")
```

```{r, eval=TRUE}
summary(meta_year)
```

\newpage

```{r, eval=TRUE, echo=FALSE}
q <- bubble_plot(meta_year, mod = "Publication.year", group="study_code")
q+ coord_cartesian(ylim=c(-20,10))
```



\newpage

# Sensitivty Analysis

Here, we perform a sensitivity analysis by removing the smallest and largest 2.5% of effect sizes.

```{r, eval=TRUE, echo=FALSE}
minq <- quantile(rdata$es, 0.025)
maxq <- quantile(rdata$es, 0.975)

sdata <- subset(rdata, es > minq & es < maxq)
```

The 2.5th and 97.5th percentiles are given below. 
```{r}
print(minq)
print(maxq)
```

The sumamry of the effect sizes is now

```{r, eval=TRUE}
summary(sdata$es)
```

We run the cubic treatment temperature and the binned temperature models again with our new subsetted data i.e. only with data that is between the 2.5Th and 97.5Th percentile.  

## Treatment temperature as a cubic effect (sesnsitivity analysis)

We re-create the variance-covariance matrix with our new subsetted data, which we name sdata. Then we run the meta analysis model again with treatment temperature as  quadratic fixed effect.

```{r, eval=TRUE}
# recreate vcv_shared matrix
VCV_shared_sa <- impute_covariance_matrix(vi=sdata$v, cluster = sdata$shared_control, r=0.5)
```

```{r, eval=FALSE, echo=TRUE}
meta_sa_treat3 <- rma.mv(es, VCV_shared_sa,  mod= ~ poly(treattemp, degree=3, raw=TRUE), random= list(~ 1|study_code,  ~1|obs), data= sdata, method= "REML")
```

```{r}
summary(meta_sa_treat3)
```


\newpage

Below is a bubble plot of the fitted cubic model to the subsetted data. 

```{r, eval=TRUE, echo=FALSE}
p <- bubble_plot(meta_sa_treat3, mod = "treattemp", group="study_code")
p+coord_cartesian(ylim=c(-10,10))
```


## Modelling response with binned treatment temperatures

Lastly, we rerun the model where our effect sizes are categorised into one seven bins depending on the treatment temperature. The bins and the number of effect sizes in each bin of the new subsetted data are given below


```{r}
table(sdata$bin.temp)
```


```{r, eval=FALSE}
meta_sa_bintemp <- rma.mv(es, VCV_shared_sa,  mod= ~bin.temp-1,  random= list(~ 1|study_code,  ~1|obs), data= sdata, method= "REML")
```

```{r, eval=TRUE}
summary(meta_sa_bintemp)
```

```{r, echo=FALSE, eval=TRUE}
plot <- orchard_plot(meta_sa_bintemp, mod = "bin.temp", xlab = "Effect size", transfm = "none", 
             cb=FALSE, group="obs", tree.order = c("<15", "15-20", "20-25", "25-30", "30-35", "35-40", ">40")) 

plot+ coord_cartesian(ylim=c(-10,10))
```


# Other fixed effects

```{r, eval=FALSE}
meta_bintemp_habitat <- rma.mv(es, VCV_shared,  mod= ~bin.temp * Habitat,  random= list(~ 1|study_code,  ~1|obs), data= rdata, method= "REML")
```

```{r}
summary(meta_bintemp_habitat)
```




